### 2023-03-02 calculus.tex
Re-marking instead of remarking
Updates from Overleaf

The key contribution of this section is the \emph{marked lambda calculus}, a calculus based on the gradually typed lambda calculus (GTLC) that formalizes the mechanism by which errors like these can be localized, and how to  recover, in all cases, from such errors. 
% In brief, the marked lambda calculus formalizes 
% bidirectional type error localization with gradual recovery. 
\Cref{sec:calculus-calculus} introduces the syntax, judgemental structure, guiding metatheory, and then goes through the rules, organized by syntactic form rather than judgement, to build intuition about how the various judgements relate to one another \cite{Siek06a}. All rules and theorems may be found organized by judgement form in the supplementary appendix for reference, 
alongside a complete mechanization in the Agda proof assistant \cite{norell:thesis} (discussed in
\cref{sec:calculus-agda}). We intentionally keep the marked lambda calculus minimal, because it is intended to capture the essential idea and introduce a general pattern that language designers can employ to created mark variants of their own type systems. As an initial example, we consider product types (in \cref{sec:calculus-products}). We then consider a surprisingly subtle combination of features as a more substantial case study: the combination of
destructing let expressions with granular type annotations, in \cref{sec:calculus-let}. Finally, \cref{sec:calculus-poly} briefly explores how
the system might be extended to more complex typing features, such as polymorphism.

@@

Alternatively, instead of a wholly untyped action calculus, one may directly integrate re-marking logic into the typed Hazelnut
calculus, making use of the type and scoping information being propagated to re-mark only in scopes where it might be necessary. For example, we can wrap $\ECMV$ into
$\ECLam{x}{\TUnknown}{\ECMV}$ by re-mark the body under the context with $x$ binding
the unknown type:
\[%
  \judgment{
    \ctxSynFixedInto{\extendCtx{\ctx}{x}{\TUnknown}}{\erase{\ECMV}}{\ECMV'}{\TMV'}
  }{
    \ASEConLam{\ctx}{\ECMV}{\TMV}{\ECMV'}{\TMV'}{x}
  }{ASEConLam}
\]%
The
supplemental material contains a complete formal description of both variants, and the untyped
action semantics and related metatheory are mechanized in Agda. Because of the simplicity of the untyped version, and because marking is defined separately, this Agda mechanization can serve universally as a baseline for correctness of approaches, like our initial design that aim to make use of type and scope information, and perhaps other analyses, to minimize re-marking. We leave a complete assessment of this and other remarking optimization approaches to future work.

>>>>

The key contribution of this section is the \emph{marked lambda calculus}, a calculus based on the gradually typed lambda calculus (GTLC) that formalizes the mechanism by which errors like these can be localized, and how to  recover, in all cases, from such errors. 
% In brief, the marked lambda calculus formalizes 
% bidirectional type error localization with gradual recovery. 
\Cref{sec:calculus-calculus} introduces the syntax, judgemental structure, guiding metatheory, and then goes through the rules, organized by syntactic form rather than judgement, to build intuition about how the various judgements relate to one another \cite{Siek06a}. All rules and theorems may be found organized by judgement form in the supplementary appendix for reference, 
alongside a complete mechanization in the Agda proof assistant \cite{norell:thesis} (discussed in
\cref{sec:calculus-agda}). We intentionally keep the marked lambda calculus minimal, because it is intended to capture the essential idea and introduce a general pattern that language designers can employ to created mark variants of their own type systems. As an initial example, we consider product types (in \cref{sec:calculus-products}). We then consider a surprisingly subtle combination of features as a more substantial case study: the combination of
destructing let expressions with granular type annotations, in \cref{sec:calculus-let}. Finally, \cref{sec:calculus-poly} briefly explores how
the system might be extended to more complex judgemental structure, e.g. polymorphism.

@@

Alternatively, instead of a wholly untyped action calculus, one may directly integrate re-marking logic into the typed Hazelnut
calculus, making use of the type and scoping information being propagated to re-mark only in scopes where it might be necessary. For example, we can wrap $\ECMV$ into
$\ECLam{x}{\TUnknown}{\ECMV}$ by re-mark the body under the context with $x$ binding
the unknown type:
\[%
  \judgment{
    \ctxSynFixedInto{\extendCtx{\ctx}{x}{\TUnknown}}{\erase{\ECMV}}{\ECMV'}{\TMV'}
  }{
    \ASEConLam{\ctx}{\ECMV}{\TMV}{\ECMV'}{\TMV'}{x}
  }{ASEConLam}
\]%
The
supplemental material contains a complete formal description of both variants, and the untyped
action semantics and related metatheory are mechanized in Agda. Because of the simplicity of the
untyped version, and because marking is defined separately, this Agda mechanization can serve
universally as a baseline for correctness of approaches, like our initial design that aim to make
use of type and scope information, and perhaps other analyses, to minimize re-marking. We leave a
complete assessment of this and other re-marking optimization approaches to future work.
